---
title: "Combine temporary files"
author: "Chang-Yu Chang"
date: "`r Sys.Date()`"
output:
  html_notebook:  
    toc: yes
    number_sections: no
---

```{r setup, include = F}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE)
library(tidyverse)
run_scripts <- F
```


# 04A-isolates

```{r}
if (run_scripts) source("script/04A-summary-01-isolates.R")
```

Isolates information in one data.frame `isolates`. 68 isolates

```{r}
isolates <- read_csv("~/Dropbox/lab/emergent-coexistence/data/output/isolates.csv", col_types = cols())
isolates
```




# 04B-pairs

```{r}
if (run_scripts) source("script/04B-summary-01-pairs.R")
```

186 pairs of self-assembled communities and 112 pair from across-community and random networks

```{r}
pairs <- read_csv("~/Dropbox/lab/emergent-coexistence/data/output/pairs.csv", col_types = cols())
pairs
```

Pairs with three initial frequencies and three. 186x3x2=1116

```{r}
pairs_freq <- read_csv("~/Dropbox/lab/emergent-coexistence/data/output/pairs_freq.csv", col_types = cols()) %>%
    filter(Set == "CFUandCASEU")
pairs_freq
```

pairs_meta contains the growth traits of isolates. Note that the order of Isolate1 and Isolate2 in some pairs are flipped such that Isolate1 is always the dominant strain and Isolate 2 is the subdomaint one. Dominant strain is the winner in exclusion pairs, and the more abundant strain (50:50 treatment) in coexistence pairs.

```{r}
if (FALSE) pairs_meta <- read_csv("~/Dropbox/lab/emergent-coexistence/data/output/pairs_meta.csv", col_types = cols())
```

Example outcome types

```{r}
if (run_scripts) source("script/04B-summary-02-outcome_types.R")
```

Coarse-grained 

```{r}
read_csv("~/Dropbox/lab/emergent-coexistence/data/output/pairs_example_outcomes.csv", col_types = cols())
```
Fine-grained 

```{r}
read_csv("~/Dropbox/lab/emergent-coexistence/data/output/pairs_example_outcomes_finer.csv", col_types = cols())
```
















