---
title: "Isolates growth rates"
author: "Chang-Yu Chang"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    number_sections: no
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE)
library(tidyverse)
library(data.table)
```

# Goal

- Read isolates growth rates

- Plot the simple fitted distribution of isolates' growth rates


# Isolate growth rates

```{r}
source("script/01D-isolates_growth_rate-01-read_data.R")
```

The isolates' growth rates are fitted from Djordje's growth curves. Duplicate measurement that were measurement on different days are removed

```{r}
isolates_growth_rate_ID
```

11 carbon sources measured

```{r}
unique(isolates_growth_rate_ID$CarbonSource)
```


Not all isolates have full 11 carbon source growth curves.

```{r}
isolates_growth_rate_ID %>%
  group_by(ExpID) %>%
  summarize(Count = n())
```


Spread the isolates community so that it contains only 68 rows (68 isolates)

```{r}
isolates_growth_rate_spread_ID
```













