---
title: "CASEU Genewix prep"
author: "Chang-Yu Chang"
date: "`r Sys.Date()`" 
output: 
    pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(data.table)
```

## Aim

Prepare the Sanger-sequencing ready sample to Genewiz `r n_plate = 1`

## Materials 

- 20 uL normalized DNA 1.25 ng/uL in clean_normalized_PCR
- PCR plate `r n_plate`
- 27F 5uM primer `r n_plate * 5 * 100 /1000` mL  
- 20uL tip box `r n_plate`

## Procedures

- [ ] Label `r n_plate` PCR plates with the naming convention **XX seq**.
- [ ] Make `r n_plate * 5 * 100` uL 5uM 27F primer stock (`r n_plate * 5 * 100 * 0.05` uL 100 uM primer stock + `r n_plate * 5 * 100 * 0.95` uL ddH2O).
- [ ] Add **5 uL** of each primer to the PCR plate seq.
- [ ] Add **10 uL** of normalized DNA to each well of the PCR plate seq.
- [ ] For <48 samples, Label tubes with tube ID from Genewiz online ordering system, which will have your initials and sample number. Write these codes on the sides of tubes. Note: For >48 samples, arrange the samples vertically (in columns).
- [ ] Store the PCR seq plates in a box. The final product is DNA with primer in a PCR plate for sequencing
- [ ] Drop off the box at Genewiz West Campus pickup location: ISTC room 366.


```{r eval = F}
data.frame(
  DNA = c("1-2ng/uL template DNA", "5uM 27F primer"),
  Volume = c(10, 5),
  FinalMass = c("20 ng", "12 pmol")
) %>%
  mutate(MasterVolume = Volume * 80) %>%
  setNames(c("DNA", "Volume (uL)", "Final Mass", "MasterVolume (uL)")) %>%
  knitr::kable()
```

