---
title: "PCR protocol for CASEU"
author: "Chang-Yu Chang"
date: "`r Sys.Date()`" 
output: 
    pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(data.table)
library(janitor)
```

## Aim

PCR the extracted DNA for CASEU `r n_plate = 1`

## Materials

-   PCR plates `r n_plate * 2`

## Procedures

### Preparation steps:

-   [ ] master mix in 96-well plate
-   [ ] PCR plates `r n_plate * 2`
-   [ ] Mastermix

```{r}
# Reagents required per well
data.frame(
  Reagent = c("ddH20", "5X HF buffer", "dNTPs (10mM)", "Phusion"),
  WellVolume = c(23.5, 10, 1, 0.5),
  stringsAsFactors = F
) %>%
  mutate(TotalVolume = WellVolume * n_plate * 108 * 2) %>%
  janitor::adorn_totals() %>% 
  knitr::kable(format = "pandoc", caption = "CASEU PCR master mix.")
```

```{r}
# Reagents used with master mix
data.frame(
  Reagent = c("PCR master mix", "27F primer (3uM)", "1492R primer (3uM)", "teamplate DNA polymerase"),
  WellVolume = c(35, 5, 5, 5), 
  stringsAsFactors = F
)  %>%
  mutate(TotalVolume = WellVolume * n_plate * 108 * 2) %>% # Duplicate PCR
#  mutate(ReservoirWellVolume = TotalVolume / 12) %>%
#  janitor::adorn_totals() %>%  # Add total volume for each well and total reagent used 
  knitr::kable(format = "pandoc", caption = "CASEU PCR reagents with master mix.")
```

-   [ ] Label `r n_plate * 2` PCR plates. For duplicated PCR.
-   [ ] Make **`r total_primer_volume = n_plate * 5 * 108 * 2; total_primer_volume` uL** of 3uM of each primer (`r total_primer_volume * 0.03` uL 100 uM stock + `r total_primer_volume * 0.97` uL ddH2O).
-   Premix the PCR reagents (total `r n_plate * 35 * 108 * 2 / 1000` mL) in a 50 mL falcon tube.
-   Use mP200 set at **35 uL** to dispense 35 uL of PCR master mix into each well of `r n_plate * 2` PCR plates. This premix can stay at room temperature.
-   Right before starting the PCR reaction, use mP20 set at **5 uL** to add primers.
-   Use mP20 set at **5 uL** to add DNA. Cover the PCR plates with clear PCR films.
-   Use the program "CASEU" in "16S" folder. See the table below for PCR cycle.
-   Store PCR plates in -20C freezer.
